---
title: "Practical Machine Learning Course Project"
author: "Sherwood Tang"
output:
  html_document:
    keep_md: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r , echo=TRUE}
library(ggplot2)
library(caret)
library(randomForest)
library(rpart)
library(rpart.plot)
```

## Loading data
```{r, echo=TRUE}
training <- read.csv("pml-training.csv", na.strings=c("NA","#DIV/0!", ""))
testing <- read.csv("pml-testing.csv", na.strings=c("NA","#DIV/0!", ""))
```


## PreProcess
# Feature Selection

First remove near zero variance features, empty columns and NA columns from test and train dataset

```{r, echo=TRUE}
nzvcol1 <- nearZeroVar(training)
training <- training[,-nzvcol1]
training <- training[,-(1:7)]
training[training == ""] <- NA
training <- training[, colSums(is.na(training)) == 0]

nzvcol2 <- nearZeroVar(testing)
testing <- testing[,-nzvcol2]
testing[testing == ""] <- NA
testing <- testing[,-(1:7)]
testing <- testing[, colSums(is.na(testing)) == 0] 
```

## Train the model


1.Partitioning
```{r , echo=TRUE}
set.seed(23333)
intrain <- createDataPartition(training$classe,p=0.7,list=FALSE)
train <- training[intrain,]
validation <- training[-intrain,]
```

2.Random forest
```{r, echo=TRUE}
modfit1 <- randomForest(classe~.,data=train,type = "class")

print(modfit1)
```

3.Rpart
```{r, echo=TRUE}
modfit2 <- rpart(classe~.,data=train)

print(modfit2)

rpart.plot(modfit2, main="Classification Tree", extra=102, under=TRUE, faclen=0)
```


4..Predict the model with validation set
```{r, echo=TRUE}
pred1 <- predict(modfit1,validation)
confusionMatrix(validation$classe, pred1)

pred2 <- predict(modfit2,validation,type="class")
confusionMatrix(validation$classe, pred2)
```

We have reached a 99.7% of accuracy by using random forest. The accuracy rate of random forest is 
higher than rpart. Therefore we choose random forest to make the prediction on test dataset.

## Apply to the test dataset

```{r, echo=TRUE}
finalpred <- predict(modfit1,testing,type="class")
finalpred
```

Save the output
```{r, echo=TRUE}
answers <- as.vector(finalpred)

pml_write_files = function(x) {
    n = length(x)
    for (i in 1:n) {
        filename = paste0("problem_id_", i, ".txt")
        write.table(x[i], file = filename, quote = FALSE, row.names = FALSE, 
            col.names = FALSE)
    }
}

pml_write_files(answers)
```